\section{Program Logics}
\label{sec:programlogics}

In this section, we describe the tactics of \EasyCrypt's three program
logics: \prhl, \phl and \hl.  There are five rough classes of program
logic tactics:
\begin{enumerate}
\item those that actually reason about the program in Hoare logic
  style;

\item those that correspond to semantics-preserving program
  transformations or compiler optimizations;

\item those that operate at the level of specifications,
  strenghtening, combining or splitting goals without modifying the
  program;

\item tactics that automate the application of other tactics;

\item advanced tactics for handling eager/lazy sampling and bounding
  the probability of failure.
\end{enumerate}
We discuss these five classes in turn.

Some of the program reasoning tactics have two modes when used on
goals whose conclusions are \prhl statement judgements.  Their default
mode is to operate on both programs at once. When a side is specified
(using \ec{$\tau$\{1\}} or \ec{$\tau$\{2\}}), a one-sided variant is
used, with 1 referring to the left program, and 2 to the right one.

\subsection{Tactics for Reasoning about Programs}
\label{subsec:reasoningprograms}

\input{tactics/proc.tex}
\input{tactics/skip.tex}
\input{tactics/seq.tex}
\input{tactics/sp.tex}
\input{tactics/wp.tex}
\input{tactics/rnd.tex}
\input{tactics/if.tex}
\input{tactics/while.tex}
\input{tactics/call.tex}

\subsection{Tactics for Transforming Programs}
\label{subsec:transformingprograms}

Unless otherwise specified, the tactics of this subsection only apply
to goals whose conclusions are \prhl, \phl or \hl statement
judgements, reducing such a goal to a single subgoal in which only the
program(s) of those statement judgements have changed.

Many of these tactics take \emph{code positions} consisting of a
sequence of positive numerals separated by dots.  E.g., \ec{2.1.3}
says to go to the statement 2 of the program, then to substatement 1
of it, then to sub-substatement 3 of it. We use the variable $c$ to
range over code positions.

\input{tactics/swap.tex}
\input{tactics/inline.tex}
\input{tactics/rcondt.tex}
\input{tactics/rcondf.tex}

\input{tactics/unroll.tex}
\input{tactics/splitwhile.tex}
\input{tactics/fission.tex}
\input{tactics/fusion.tex}

\input{tactics/alias.tex}
\input{tactics/cfold.tex}
\input{tactics/kill.tex}

\subsection{Tactics for Reasoning about Specifications}

\input{tactics/symmetry.tex}
\input{tactics/transitivity.tex}
\input{tactics/conseq.tex}
\input{tactics/case_pl.tex}
\input{tactics/byequiv.tex}
\input{tactics/byphoare.tex}
\input{tactics/hoare.tex}
\input{tactics/bypr.tex}
\input{tactics/exists_star.tex}
\input{tactics/elim_star.tex}
\input{tactics/exfalso.tex}
\input{tactics/phoare_split.tex}

\subsection{Automated Tactics}
\label{subsec:automatedtactics}

\input{tactics/auto.tex}
\input{tactics/sim.tex}

\subsection{Advanced Tactics}
\label{subsec:advancedtactics}

\input{tactics/eager.tex}
\input{tactics/fel.tex}
