\chapter{Tactics}
\label{chap:tactics}

Proofs in \EasyCrypt are carried out using \emph{tactics}, logical
rules embodying general reasoning principles, which transform the
current lemma (or \emph{goal}) into zero or more
\emph{subgoals}---sufficient conditions for the lemma/goal to
hold. Simple ambient logic goals may be automatically proved using SMT
solvers.

In this chapter, we introduce \EasyCrypt's proof engine, before
describing the tactics for \EasyCrypt's four logics: ambient, \prhl,
\phl and \hl.

\section{Proof Engine}

\EasyCrypt's proof engine works with goal lists, where a \emph{goal}
has two parts:
\begin{itemize}
\item A \emph{context} consisting of a
  \begin{itemize}
  \item a set of type variables, and

  \item an \emph{ordered} set of \emph{assumptions}, consisting of
    identifiers with their types, memories, module names with their
    module types and restrictions, local definitions, and named
    formulas. An identifier's type may involve the type variables, the
    local definitions and formulas may involve the type variables,
    identifiers, memories and module names.
  \end{itemize}

\item A \emph{conclusion}, consisting of a single formula, with
  the same constraints as the assumption formulas.
\end{itemize}
Informally, to prove a goal, one must show the conclusion to be true,
given the truth of the formula assumptions, for all valid instantiations
of the assumption identifiers, memories and module names.

For example,
\ecinput{examps/parts/tactics-examp0-2.1.ec}{}{}{}{}
is a goal.
And, in the context of the declarations
\ecinput{examps/tactics-examp2.ec}{}{3-11}{}{}
this is a goal:
\ecinput{examps/parts/tactics-examp2-2.0.ec}{}{}{}{}
The conclusion of this goal is just a nonlinear rendering of the formula
\begin{easycrypt}{}{}
phoare [G(X).g : G.x = n ==> G.x = n] = 1%r.
\end{easycrypt}
\EasyCrypt's pretty printer renders \prhl, \phl and \hl judgements
in such a nonlinear style when the judgements appear as
(as opposed to in) the conclusions of goals.

Internally, \EasyCrypt's proof engine also works with \prhl, \phl and
\hl judgments involving lists of statements rather than procedure
names, which we'll call \emph{statement judgements}, below. For
example, given this declaration
\ecinput{examps/tactics-examp1.ec}{}{3-9}{}{}
this is an \phl statement judgement:
\ecinput{examps/parts/tactics-examp1-3.0.ec}{}{}{}{}
The pre- and post-conditions of a statement judgement may refer to the
parameters and local variables of the procedure that's the
\emph{context} of the conclusion---\ec{M.f} in the preceding
example. They may also refer to the memories \ec{&$1$} and \ec{&$2$}
in the case of \prhl statement judgements.
When a statement judgement appears anywhere other than as the conclusion of
a goal, the pretty printer renders it in abbreviated linear syntax.
E.g., the preceding goal is rendered as
\begin{easycrypt}{}{}
hoare[if (x %% 3 = 1) {...} : x %% 3 = n ==> x %% 3 = n %% 2 + 1]
\end{easycrypt}
Statement judgements can't be directly input by the user.

When the proof of a lemma is begun, the proof engine starts out with
a single goal, consisting of the lemma's statement. E.g.,
the lemma
\ecinput{examps/tactics-examp0.ec}{}{1-3}{}{}
gives rise to the goal
\ecinput{examps/parts/tactics-examp0-1.0.ec}{}{}{}{}
For parameterized lemmas, the goal includes the lemma's parameters
as assumptions. E.g.,
\begin{easycrypt}{}{}
lemma PairEq (x x' : 'a) (y y' : 'b) :
  x = x' => y = y' => (x, y) = (x', y').
\end{easycrypt}
gives rise to
\ecinput{examps/parts/tactics-examp0-2.0.ec}{}{}{}{}

\EasyCrypt's tactics, when applicable, reduce the first goal to zero
or more subgoals.  E.g., if the first goal is
\ecinput{examps/parts/tactics-examp1-3.0.ec}{}{}{}{}
then applying the \ec{if} tactic (handle a conditional) reduces
(replaces) this goal with the two goals
\ecinput{examps/parts/tactics-examp1-3.1.ec}{}{}{}{} and
\ecinput{examps/parts/tactics-examp1-3.2.ec}{}{}{}{}
(leaving the remaining goals, if any, unchanged).
If the first goal is
\ecinput{examps/parts/tactics-examp1-6.0.ec}{}{}{}{}
then applying the \ec{smt} tactic (try to solve the goal using
SMT provers) solves the goal, i.e., replaces it with no subgoals.
Applying a tactic may fail; in this case an error message is issued
and the list of goals is left unchanged.

A lemma's proof may be saved, using the step \ec{qed}, when the list
of goals becomes empty. And this must be done before anything else may
be done.

\section{Ambient logic}
\label{sec:ambientlogic}

In this section, we describe the proof terms, tactics and tacticals of
\EasyCrypt's ambient logic.

\subsection{Proof Terms}
\label{subsec:proofterms}

Formulas introduce identifier and formula assumptions using universal
quantifiers and implications. For example, the formula
\begin{easycrypt}{}{}
forall (x y : bool), x = y => forall (z : bool), y = z => x = z.
\end{easycrypt}
introduces the assumptions
\begin{easycrypt}{}{}
x     : bool
y     : bool
eq_xy : x = y
z     : bool
eq_yz : y = z
\end{easycrypt}
(where the names of the two formulas were chosen to be meaningful),
and has \ec{x = z} as its conclusion. We refer to the first assumption
of a formula as the formula's \emph{top assumption}. E.g., the top
assumption of the preceding formula is \ec{x : bool}.

\EasyCrypt has \emph{proof terms}, which partially describe how
to prove a formula.  Their syntax is described in Figure~\ref{fig:proofterms},
where $X$ ranges over lemma (or formula assumption) names.
\begin{figure}
  \begin{center}
  \begin{tabular}{rcl>{\bf}l}
    $p$ & ::=
      & {\ec{_}} & proof hole \\
     && {\ec{($X$, $\;q_1$, $\;\ldots$, $\;q_n$)}} & lemma application \\[.2cm]
    $q$ & ::=
      & {$e$} & expression \\
      && {$p$} & proof term \\
  \end{tabular}
  \end{center}
  \caption{\label{fig:proofterms}Proof Terms}
\end{figure}
A proof term for a lemma (or formula assumption) $X$ has components
corresponding to the assumptions introduced by $X$.  A component
corresponding to a variable consists of an expression of the
variable's type. The proof term is explaining how the instantiation of
the lemma's conclusion with these expressions may be proved.  A
formula component consists of a proof term explaining how the
instantiation of the formula may be proved.  Proof holes will get
turned into subgoals when a proof term is used in backward reasoning,
e.g., by the \rtactic{apply} tactic. \fix{Need explanation of how a
  proof term may be used in forward reasoning.}

Consider, e.g., the following declarations and axioms
\begin{easycrypt}{}{}
pred P : int.
pred Q : int.
pred R : int.
axiom P (x : int) : P x.
axiom Q (x : int) : P x => Q x.
axiom R (x : int) : P(x + 1) => Q x => R x.
\end{easycrypt}
Then, given that \ec{x : int} is an assumption,
\begin{easycrypt}{}{}
(R x (P(x + 1)) (Q x (P x)))
\end{easycrypt}
is a proof term proving the conclusion \ec{R x}. And
\begin{easycrypt}{}{}
(R x _ (Q x _))
\end{easycrypt}
is a proof term that turns proofs of \ec{P(x + 1)} and \ec{P x}
into proofs of \ec{R x}. When used in backward reasoning, it
will reduce a goal with conclusion \ec{R x} to subgoals with
conclusions \ec{P(x + 1)} and \ec{P x}.
\fix{Can it be used in forward reasoning?}

Some of a proof term's expressions may be replaced by \ec{_}, asking
\EasyCrypt to infer them from the context.  Going even further, one
may abbreviate a one-level proof term all of whose components are
\ec{_} to just its lemma name. For example, we can write \ec{R} for
\ec{(R _ _ _)}.  When used in backward reasoning, it will reduce a
goal with conclusion \ec{R x} to subgoals with conclusions \ec{P(x +
  1)} and \ec{Q x}. \fix{In forward reasoning they aren't equivalent---why?}

\subsection{Occurrence Selectors and Rewriting Directions}
\label{subsec:occsels}

Some ambient logic tactics use \emph{occurrence selectors} to restrict
their operation to certain occurrences of a term or formula in a
goal's conclusion or formula assumption. The syntax is \ec{\{$i_1$,
  $\;\ldots$, $\;i_n$\}}, specifying that only occurrences $i_1$
throught $i_n$ of the term/formula in a depth-first, left-to-right
traversal of the goal's conclusion or formula assumption should be
operated on. Specifying \ec{\{- $i_1$, $\;\ldots$, $\;i_n$\}}
restricts attention to all occurrences \emph{not} in the following
list. They may also be empty, meaning that all applicable occurrences
should be operated on.

Some ambient logic tactics uses \emph{rewriting directions}, $\mathit{dir}$,
which may either be empty (meaning rewriting from left to right), or \ec{-},
meaning rewriting from right to left.

\subsection{Introduction and Generalization}
\label{subsec:introgen}

One moves the assumptions of a goal's conclusion into the goal's
context using the introduction tactical.  This tactical uses
introduction patterns, which are defined in Figure~\ref{fig:intropat}.
\begin{figure}
  \begin{center}
  \begin{tabular}{rcl>{\bf}l}
    $\iota$ & ::=
      & {$b$} & name \\
      && {\ec{$b\,$!}} & rename \\
      && {\ec{_}} & no name \\
      && {\ec{+}} & auto revert \\
      && {\ec{?}} & find name \\
      && {$\mathit{occ}\;$\ec{->}} & rewrite using assumption \\
      && {$\mathit{occ}\;$\ec{<-}} & rewrite in reverse using assumption \\
      && {\ec{->>}} & substitute using assumption \\
      && {\ec{<<-}} & substitute in reverse using assumption \\
      && {\ec{/$p$}} & replace assumption by applying proof term \\
      && {\ec{\{$b_1\cdots b_n$\}}} & clear introduced assumptions \\
      && {\ec{/=}} & simplify \\
      && {\ec{//}} & trivial \\
      && {\ec{//=}} & simplify then trivial \\
      && {\ec{$\mathit{dir}\;\mathit{occ}\;$@/$\mathit{op}$}} & unfold definition of operator \\
      && {\ec{[$\iota_{11}\cdots\iota_{1{m_1}}$ | $\;\cdots$ | $\;\iota_{r1}\cdots\iota_{r{m_r}}$]}} & case pattern \\[.2cm]
    $\mathit{b}$ & ::=
      & {$x$} & identifier \\
      && {$M$} & module name \\
      && {\ec{&$m$}} & memory name \\
  \end{tabular}
  \end{center}
  \caption{\label{fig:intropat} Introduction Patterns}
\end{figure}
In this definition, $\mathit{occ}$ ranges over occurrence selectors,
and $\mathit{dir}$ ranges over directions---see
Subsection~\ref{subsec:occsels}).

If a list $\iota_1,\ldots,\iota_n$ of introduction patterns consists
entirely of \ec{//} (apply the \rtactic{trivial} tactic), \ec{/=}
(apply the \rtactic{simplify} tactic) and \ec{//=} (apply the
\ec{simplify} and then \ec{trivial}), then \emph{applying}
$\iota_1,\ldots,\iota_n$ \emph{to} a list of goals $G_1,\ldots,G_m$ is
done by applying the tactics corresponding to the $\iota_i$ in order
to each $G_j$, causing some of the goals to be solved and thus
disappear and some of the goals to be simplified.

\begin{tactic}[$\tau$=>$\;\iota_1 \cdots \iota_n$]{introduction}
  \begin{tsyntax}[empty]{}
    Runs the tactic $\tau$, matching the resulting goals, $G_1,\ldots,G_l$,
    with the introduction patterns $\iota_1,\ldots,\iota_n$:
    \begin{itemize}
    \item Suppose $k$ is such that all of $\iota_1,\ldots,\iota_{k-1}$
      are \ec{//}, \ec{/=} and \ec{//=}, and either $k>n$ or $\iota_k$
      is not \ec{//}, \ec{/=} or \ec{//=}.

    \item Let $G'_1,\ldots,G'_{l'}$ be the goals resulting from
      applying $\iota_1,\ldots,\iota_{k-1}$ to $G_1,\ldots,G_l$.

    \item If $l'=0$, the tactical produces no subgoals.

    \item Otherwise, if $k>n$, the tactical's result is
      $G'_1,\ldots,G'_{l'}$.

    \item Otherwise, if $\iota_k$ is not a case pattern, each subgoal
      $G'_i$ is matched against $\iota_k,\ldots,\iota_n$ by the
      procedure described below, with the resulting subgoals being
      collected into a list of goals (maintaining order viz a viz the
      indices $i$) as the tactical's result.

    \item Otherwise, $\iota_k$ is a case pattern
          \ec{[$\iota_{11}\cdots\iota_{1{m_1}}$ | $\;\cdots$ | $\;\iota_{r1}\cdots\iota_{r{m_r}}$]}.

     \item If $\tau$ is not equivalent to \rtactic{idtac}, the tactic fails
       unless $r = l'$, in which case each $G'_i$ is matched against 
        \begin{displaymath}
          \iota_{i1}\cdots\iota_{i{m_i}}\iota_{k+1}\cdots \iota_n
        \end{displaymath}
       by the procedure described below, with the resulting subgoals
       being collected into the tactical's result.

     \item Otherwise, $\tau$ is equivalent to \rtactic{idtac} (and so
       $l'=1$). In this case $G'_1$ is matched against
       $\iota_k,\ldots,\iota_n$ by the procedure described below, with
       the resulting subgoals being collected into a list of goals as
       the tactical's result.
    \end{itemize}

    \paragraph{Matching a single goal against a list of patterns:}

    To match a goal $G$ against a list of introduction patterns
    $\iota_1,\ldots,\iota_n$, the introduction patterns are processed
    from left-to-right, as follows:
    \begin{itemize}
    \item ($b$)\quad The top assumption (universally quantified
      identifier, module name or memory; or left side of implication)
      is consumed, and introduced with this name. Fails if the top
      assumption has neither of these forms.

    \item (\ec{$b\,$!})\quad Same as the preceding case, except that
      $b$ is used as the base of the introduced name, extending the
      base to avoid naming conflicts.

    \item (\ec{_})\quad Same as the preceding case, except the
      assumption is introduced with an anonymous name (which can't be
      uttered by the user).

    \item (\ec{+})\quad Same as the preceding case, except that after
      a branch of the procedure completes, yielding a goal, the
      assumption will be reverted, i.e., un-introduced (using a
      universal quantifier or implication as appropriate).

    \item (\ec{?})\quad Same as the preceding case, except \EasyCrypt
      chooses the name by which the assumption is introduced (using
      universally quantified names as assumption bases).

    \item ($\mathit{occ}\;$\ec{->})\quad Consume the top assumption,
      which must be an equality, and use it as a left-to-right rewriting
      rule in the remainder of the goal's conclusion, restricting rewriting
      to the specified occurrences of the equality's left side.

    \item ($\mathit{occ}\;$\ec{<-})\quad The same as the preceding case,
      except the rewriting is from right-to-left.

    \item (\ec{->>})\quad The same as \ec{->}, except the consumed
      equality assuption is used to perform a left-to-right substitution
      in the entire goal, i.e., in its assumptions, as well as its
      conclusion.

    \item (\ec{<<-})\quad The same as the preceding case, except
      the substitution is from right-to-left.

    \item (\ec{/$p$})\quad Replace the top assumption by the result
    of applying the proof term $p$ to it using forward reasoning.

    \item (\ec{\{$b_1\cdots b_n$\}})\quad Doesn't affect the goal's
      conclusion, but clears the specified assumptions, i.e., removes
      them. Fails if one or more of the assumptions can't be cleared,
      because a remaining assumption depends upon it.

    \item (\ec{/=})\quad Apply \rtactic{simplify} to goal's conclusion.

    \item (\ec{//})\quad Apply \rtactic{trivial} to goal's conclusion;
      this may solve the goal, i.e., so that the procedure's current
      branch yields no goals.

    \item (\ec{/=})\quad Apply \rtactic{simplify} and then \rtactic{trivial}
      to goal's conclusion; this may solve the goal, so that the
      procedure's current branch yields no goals.

    \item ({\ec{$\mathit{dir}\;\mathit{occ}\;$@/$\mathit{op}$}})\quad
      Unfold (fold, if the direction is \ec{-}) the definition of
      operator $\mathit{op}$ at the specified occurrences of the
      goal's conclusion.

    \item (\ec{[$\iota_{11}\cdots\iota_{1{m_1}}$ | $\;\cdots$ | $\;\iota_{r1}\cdots\iota_{r{m_r}}$]})\quad
      \begin{itemize}
      \item If $r=0$, then the top assumption of the goal is destructed
        using the \rtactic{case} tactic, the resulting goals are
        matched against $\iota_2,\ldots,\iota_n$, and their subgoals
        are assembled into a list of goals.

      \item Otherwise $r>0$. The goal's top assumption is destructed
        using the \rtactic{case} tactic, yielding subgoals
        $H_1,\ldots H_p$.  If $p\neq r$, the procedure fails. Otherwise
        each subgoal $H_i$ is matched against
        \begin{displaymath}
          \iota_{i1}\cdots\iota_{i{m_i}}\iota_2\cdots \iota_n
        \end{displaymath}
        with the resulting goals being collected into a list as
        the procedure's result.
      \end{itemize}
    \end{itemize}

    The following examples use the tactic \rtactic{move}, which is
    equivalent to \rtactic{idtac}.
    In its simplest form, the introduction tactical simply gives names
    to assumptions.  For example, if the current goal is
    \ecinput{examps/parts/tactics/introduction/1-1.0.ec}{}{}{}{}
    then running
    \ecinput{examps/parts/tactics/introduction/1-1.ec}{}{}{}{}
    produces
    \ecinput{examps/parts/tactics/introduction/1-1.1.ec}{}{}{}{}
    Alternatively, we can use the introduction pattern \ec{?}
    to let \EasyCrypt choose the assumption names, using
    \ec{H} as a base for formula assumptions and starting
    from the identifier names given in universal quantifiers:
    \ecinput{examps/parts/tactics/introduction/2-1.ec}{}{}{}{}
    produces
    \ecinput{examps/parts/tactics/introduction/2-1.1.ec}{}{}{}{}
    Or we can use the \ec{!} pattern suffix to specify our
    own base assumption names: running
    \ecinput{examps/parts/tactics/introduction/3-1.ec}{}{}{}{}
    produces
    \ecinput{examps/parts/tactics/introduction/3-1.1.ec}{}{}{}{}

    To see how the \ec{->} rewriting pattern works, suppose
    the current goal is
    \ecinput{examps/parts/tactics/introduction/4-1.0.ec}{}{}{}{}
    Then running
    \ecinput{examps/parts/tactics/introduction/4-1.ec}{}{}{}{}
    produces
    \ecinput{examps/parts/tactics/introduction/4-1.1.ec}{}{}{}{}
    Alternatively, one can introduce the assumption \ec{x = y},
    and then use the \ec{->>} substitution pattern:
    if the current goal is
    \ecinput{examps/parts/tactics/introduction/8-1.0.ec}{}{}{}{}
    then running
    \ecinput{examps/parts/tactics/introduction/8-1.ec}{}{}{}{}
    produces
    \ecinput{examps/parts/tactics/introduction/8-1.1.ec}{}{}{}{}

    To see how a view may be applied to a not-yet-introduced formula
    assumption, suppose the current goal is
    \ecinput{examps/parts/tactics/introduction/5-1.0.ec}{}{}{}{}
    Then running
    \ecinput{examps/parts/tactics/introduction/5-1.ec}{}{}{}{}
    produces
    \ecinput{examps/parts/tactics/introduction/5-1.1.ec}{}{}{}{}
    And then running
    \ecinput{examps/parts/tactics/introduction/5-2.ec}{}{}{}{}
    on this goal produces
    \ecinput{examps/parts/tactics/introduction/5-2.1.ec}{}{}{}{}

    Finally, let's see examples of how a disjunction assumption
    may be destructed, either using the \ec{case} tactic followed
    by a case introduction pattern, or by making the
    case introduction pattern do the destruction.
    For the first case, if the current goal is
    \ecinput{examps/parts/tactics/introduction/6-1.0.ec}{}{}{}{}
    then running
    \ecinput{examps/parts/tactics/introduction/6-1.ec}{}{}{}{}
    produces the two goals
    \ecinput{examps/parts/tactics/introduction/6-1.1.ec}{}{}{}{}
    and
    \ecinput{examps/parts/tactics/introduction/6-1.2.ec}{}{}{}{}
    And for the second case, if the current goal is
    \ecinput{examps/parts/tactics/introduction/7-1.0.ec}{}{}{}{}
    then running
    \ecinput{examps/parts/tactics/introduction/7-1.ec}{}{}{}{}
    produces the two goals
    \ecinput{examps/parts/tactics/introduction/7-1.1.ec}{}{}{}{}
    and
    \ecinput{examps/parts/tactics/introduction/7-1.2.ec}{}{}{}{}
    Note how we used the clear pattern to discard the assumption
    \ec{X}.
  \end{tsyntax}
\end{tactic}

Generalization moves assumptions back into a goal's conclusion:

\begin{tactic}[$\tau$: $\;\pi_1 \cdots \pi_n$]{generalization}
  \begin{tsyntax}[empty]{$\tau$: $\;\pi_1 \cdots \pi_n$}
    Generalize the patterns $\pi_1, \cdots, \pi_n$, starting from
    $\pi_n$ and going back, and then run tactic $\tau$.
    This tactical is only applicable to certain tactics:
    \rtactic{move}, \rtactic{case} and \rtactic{elim}.

    \begin{itemize}
    \item If a pattern $\pi$ is an assumption, it's moved back into
      the conclusion, using universal quantification or an
      implication, as appropriate. If one assumption depends on
      another, one can't generalize the later without also
      generalizing the former.

      For example, if the current goal is
      \ecinput{examps/parts/tactics/generalize/2-1.0.ec}{}{}{}{} then
      running \ecinput{examps/parts/tactics/generalize/2-1.ec}{}{}{}{}
      produces
      \ecinput{examps/parts/tactics/generalize/2-1.1.ec}{}{}{}{} In
      this example, one can't generalize \ec{x} without also
      generalizing \ec{eq_xy}.

    \item A pattern $\pi$ may also be a subformula or subterm of the
      goal, or \ec{_}, which stands for the whole goal, possibly
      prefixed by an occurrence selector. This replaces the formula or
      subterm with a universally quantified identifier of the
      approprate type.

      For example, if the current goal is
      \ecinput{examps/parts/tactics/generalize/1-1.0.ec}{}{}{}{} then
      running \ecinput{examps/parts/tactics/generalize/1-1.ec}{}{}{}{}
      produces
      \ecinput{examps/parts/tactics/generalize/1-1.1.ec}{}{}{}{}
      Alternatively, running
      \ecinput{examps/parts/tactics/generalize/3-1.ec}{}{}{}{}
      produces
      \ecinput{examps/parts/tactics/generalize/3-1.1.ec}{}{}{}{}
    \end{itemize}
  \end{tsyntax}
\end{tactic}

\subsection{Tactics}

\input{tactics/idtac.tex}
\input{tactics/move.tex}
\input{tactics/clear.tex}
\input{tactics/assumption.tex}
\input{tactics/reflexivity.tex}
\input{tactics/left.tex}
\input{tactics/right.tex}
\input{tactics/split.tex}
\input{tactics/congr.tex}
\input{tactics/subst.tex}

\input{tactics/trivial.tex}
\input{tactics/done.tex}
\input{tactics/simplify.tex}
\input{tactics/progress.tex}
\input{tactics/smt.tex}
\input{tactics/admit.tex}

\input{tactics/change.tex}
\input{tactics/pose.tex}
\input{tactics/have.tex}
\input{tactics/cut.tex}

\input{tactics/apply.tex}
\input{tactics/exact.tex}
\input{tactics/rewrite.tex}

\input{tactics/case.tex}
\input{tactics/elim.tex}

\input{tactics/algebra.tex}

\subsection{Tacticals}
\label{subsec:tacticals}

Tactics can be combined together, composed and modified by
\emph{tacticals}. We've already seen the introduction and generalization
tacticals, which turn a tactic $\tau$ and a list of patterns into
a composite tactic, which may then combined with other tactics.

\begin{tactic}[$\tau_1$; $\;\tau_2$]{sequence}
  \begin{tsyntax}[empty]{t1; t2}
  Apply $\tau_2$ to all the subgoals generated by $\tau_1$.

  \paragraph{Remark.} Sequencing groups to the left, so that
    \ec{$\tau_1$; $\;\tau_2$; $\;\tau_3$} means
    \ec{($\tau_1$; $\;\tau_2$); $\;\tau_3$}.
  \end{tsyntax}
\end{tactic}

\begin{tactic}[try $\;\tau$]{failure recovery}\label{tactic-try}
  \begin{tsyntax}[empty]{try t}
  Execute the tactic $\tau$ if it succeeds; do nothing (leave the
  goal unchanged) if it fails.

  \paragraph{Remark.}
  By default, \EasyCrypt proofs are run in \ec{strict} mode. In this
  mode, \ec{smt} failures cannot be caught using \ec{try}. This allows
  \EasyCrypt to always build the proof tree correctly, even in weak
  check mode, where \ec{smt} calls are assumed to succeed. Inside a
  strict proof, weak check mode can be turned on and off at will,
  allowing for the fast replay of proof sections during
  development. In any event, we recommend \emph{never} using \ec{try
    smt}: a little thought is much more cost-effective than failing
  \ec{smt} calls.
  \end{tsyntax}
\end{tactic}

\begin{tactic}[do! $\;\tau$]{tactic repetition}
  \begin{tsyntax}[empty]{do! t}
    Apply $\tau$ to the current goal, then repeatedly apply it to all
    subgoals, stopping on a branch only when it fails. An error is
    produced it $\tau$ does not apply to the current goal.
  \end{tsyntax}

  \paragraph{Variants.}\strut

  \begin{tabularx}{\textwidth}{@{}ll@{}}
  {\ec{do? $\;\tau$}} & apply $\tau$ 0 or more times, until it fails\\
  {\ec{do $\;n$! $\;\tau$}} & apply $\tau$ with depth exactly $n$\\
  {\ec{do $\;n$? $\;\tau$}} & apply $\tau$ with depth at most $n$
  \end{tabularx}
\end{tactic}

\begin{tactic}[$\tau$; first $\;\tau_2$]{goal selection}
  \begin{tsyntax}[empty]{t1; first t2}
    Apply the tactic $\tau_1$, then apply $\tau_2$ on the first
    subgoal generated by \ec{t1}, leaving the other goals unchanged.
    An error is produced if no subgoals are generated by $\tau_1$.

  \paragraph{Variants.}\strut

  \noindent\begin{tabularx}{\textwidth}{@{}ll@{}}
    {\ec{$\tau_1$; first $\;n$ $\;\tau_2$}} & apply $\tau_2$ on the
    first $n$ subgoals generated by $\tau_1$\\[.4cm]
    {\ec{$\tau_1$; last $\;\tau_2$}} & apply $\tau_2$ on the last subgoal
    generated by $\tau_1$\\[.4cm]
    {\ec{$\tau_1$; last $\;n$ $\;\tau_2$}} & apply $\tau_2$ on the last $n$
    subgoals generated by $\tau_1$\\[.4cm]
    {\ec{$\tau$; first $\;n\!$ last}} & \parbox{250pt}{reorder the subgoals
      generated by $\tau$, moving the first $n$ to the end of the
      list} \\[.4cm]
    {\ec{$\tau$; last $\;n\!$ first}} & \parbox{250pt}{reorder the subgoals
      generated by $\tau$, moving the last $n$ to the beginning of the
      list} \\[.4cm]
    {\ec{$\tau$; last first}} & \parbox{250pt}{reorder the subgoals generated
    by $\tau$, moving the last one to the beginning of the list}\\[.4cm]
    {\ec{$\tau$; first last}} & \parbox{250pt}{reorder the subgoals
     generated by $\tau$, moving the first one to the end of the list}
  \end{tabularx}
  \end{tsyntax}
\end{tactic}

\begin{tactic}[by $\;\tau$]{closing goals}
  \begin{tsyntax}[empty]{by t}
  Apply the tactic $\tau$ and try to close all the generated subgoals using
  \rtactic{trivial}. Fail if not all subgoals can be closed.
  \end{tsyntax}

  \paragraph{Remark.} Inside the a lemma's proof, \ec{by []} is
  equivalent to \ec{by trivial}.  But the form
\begin{easycrypt}{}{}
lemma #$\;\cdots$# by [].
\end{easycrypt}
  means
\begin{easycrypt}{}{}
lemma #$\;\cdots$# by (trivial; smt).
\end{easycrypt}
\end{tactic}

\section{Program Logics}
\label{sec:programlogics}

In this section, we describe the tactics of \EasyCrypt's three program
logics: \prhl, \phl and \hl.  There are five rough classes of program
logic tactics:
\begin{enumerate}
\item those that actually reason about the program in Hoare logic
  style;

\item those that correspond to semantics-preserving program
  transformations or compiler optimizations;

\item those that operate at the level of specifications,
  strenghtening, combining or splitting goals without modifying the
  program;

\item tactics that automate the application of other tactics;

\item advanced tactics for handling eager/lazy sampling and bounding
  the probability of failure.
\end{enumerate}
We discuss these five classes in turn.

\subsection{Tactics for Reasoning about Programs}
\label{subsec:reasoningprograms}

Unless specified, the following program logic tactics operate on a
program's last statement. Although we describe these tactics as if
they operated on single statements, their practical implementation
automatically and implicitly applies tactic \rtactic{seq} to deal with
context when necessary.

For simple proofs, it is often enough to simply apply the program
tactic corresponding to the last statement in the program and let
\ec{smt} deal with the residual program-free formula, once the program
has been consumed.

Most of the program reasoning tactics discussed in this subsection
have two modes when used on \prhl proof obligations. Their default
mode is to operate on both programs at once. When a side is specified
(using \ec{<tactic>\{1\}} or \ec{<tactic>\{2\}}), a one-sided variant
is used. Apart from the \rtactic{if} tactic, the one-sided variant is
in fact a combination of the \phl tactic and \rtactic{conseq}.

\medskip

\input{tactics/skip.tex}
\input{tactics/seq.tex}
\input{tactics/sp.tex}
\input{tactics/wp.tex}
\input{tactics/rnd.tex}
\input{tactics/if.tex}
\input{tactics/while.tex}
\input{tactics/call.tex}
\input{tactics/proc.tex}

\subsection{Tactics for Transforming Programs}
\label{subsec:transformingprograms}

\input{tactics/swap.tex}
\input{tactics/inline.tex}

\input{tactics/rcondf.tex}
\input{tactics/rcondt.tex}

\input{tactics/splitwhile.tex}
\input{tactics/unroll.tex}
\input{tactics/fission.tex}
\input{tactics/fusion.tex}

\input{tactics/alias.tex}
\input{tactics/cfold.tex}
\input{tactics/kill.tex}
\input{tactics/modpath.tex}

\subsection{Tactics for Reasoning about Specifications}

\input{tactics/symmetry.tex}
\input{tactics/transitivity.tex}
\input{tactics/conseq.tex}
\input{tactics/case_pl.tex}
\input{tactics/phoare_split.tex}
\input{tactics/byequiv.tex}
\input{tactics/byphoare.tex}
\input{tactics/hoare.tex}
\input{tactics/bypr.tex}
\input{tactics/exists_star.tex}
\input{tactics/elim_star.tex}
\input{tactics/exfalso.tex}

\subsection{Automated Tactics}
\label{subsec:automatedtactics}

\input{tactics/auto.tex}
\input{tactics/sim.tex}

\subsection{Advanced Tactics}
\label{subsec:advancedtactics}

\input{tactics/eager.tex}
\input{tactics/fel.tex}
