% --------------------------------------------------------------------
\begin{tactic}[apply $\;p$ | apply /$p_1$ $\;\cdots$ /$p_n$ | apply $\;p$ in $\;H$]{apply}
  \begin{tsyntax}{apply $\;p$}
    Tries to match the conclusion of the proof term $p$ with the
    goal's conclusion. If the match succeeds and leads to the full
    instantiation of the pattern, then the goal is replaced, after
    instantiation, with the subgoals of the proof term.

    Consider the declarations
\begin{easycrypt}{}{}
pred P : int.
pred Q : int.
pred R : int.
axiom P (x : int) : P x.
axiom Q (x : int) : P x => Q x.
axiom R (x : int) : P(x + 1) => Q x => R x.
\end{easycrypt}
    If the current goal is
    \ecinput{examps/parts/tactics/apply/1-1.0.ec}{}{}{}{} then
    running \ecinput{examps/parts/tactics/apply/1-1.ec}{}{}{}{}
    produces the goals
    \ecinput{examps/parts/tactics/apply/1-1.1.ec}{}{}{}{}
    and
    \ecinput{examps/parts/tactics/apply/1-1.2.ec}{}{}{}{}
    And running
    \ecinput{examps/parts/tactics/apply/2-1.ec}{}{}{}{}
    from that initial goal
    produces the goals
    \ecinput{examps/parts/tactics/apply/2-1.1.ec}{}{}{}{}
    and
    \ecinput{examps/parts/tactics/apply/2-1.2.ec}{}{}{}{}
  \end{tsyntax}

  \begin{tsyntax}{apply /$p_1$ $\;\cdots$ /$p_n$}
    Apply the proof terms $p_1,\ldots,p_n$ in sequence. At each stage
    of this process, we have some number of goals. Initially, we have
    just the current goal. After applying $p_1$, we have whatever
    goals $p_1$ has produced from the current goal. $p_2$ is applied
    to the \emph{last} of these goals, and that last goal is replaced
    by the goals produced by running $p_2$, etc.  Fails without
    changing the goal if any of these applications fails.

    For example, if the current goal is
    \ecinput{examps/parts/tactics/apply/3-1.0.ec}{}{}{}{} then
    running \ecinput{examps/parts/tactics/apply/3-1.ec}{}{}{}{}
    solves the goal.
  \end{tsyntax}

  \begin{tsyntax}{apply $\;p$ in $\;H$}
    Apply $p$ in forward reasoning to $H$, replacing $H$ by the result.

    For example, if the current goal is
    \ecinput{examps/parts/tactics/apply/4-1.0.ec}{}{}{}{} then
    running \ecinput{examps/parts/tactics/apply/4-1.ec}{}{}{}{}
    produces the goal
    \ecinput{examps/parts/tactics/apply/4-1.1.ec}{}{}{}{}
  \end{tsyntax}
\end{tactic}
