% --------------------------------------------------------------------
\begin{tactic}{swap}
All versions of the tactic work for \prhl (an optional side can be given),
\phl and \hl statement judgements. We'll describe their operation
in terms of a single program (list of statements).

\medskip
\begin{tsyntax}{swap $\;n$ $\;m$ $\;l$}
  Fails unless $1\leq n < m \leq l$ and the program has at least $l$
  statements. Swaps the statement block from positions $n$ through
  $m-1$ with the statement block from $m$ through $l$, failing if these
  blocks of statements aren't independent.
\end{tsyntax}

\begin{tsyntax}{swap [$n$..$m$] $\;k$}
  Fails unless $1\leq n \leq m$ and the program has at
  least $m$ statements.
  \begin{itemize}
  \item If $k$ is non-negative, move the statement block from $n$
    through $m$ forward $k$ positions, failing if the program doesn't
    have at least $m + k$ statements or if the swapped statements
    blocks aren't independent.

  \item If $k$ is negative, move the statement block from $n$ through
    $m$ backward $-k$ positions, failing if $n + k < 1$ or if the
    swapped statement blocks aren't independent.
  \end{itemize}
\end{tsyntax}

\begin{tsyntax}{swap $\;n$\ $k$}
  Equivalent to \ec{swap [$n$..$n$] $\;k$}.
\end{tsyntax}

\begin{tsyntax}{swap $\;k$}
  If $k$ is non-negative, equivalent to \ec{swap 1 $\;k$}.
  If $k$ is negative, equivalent to \ec{swap $\;n$ $\;k$},
  where $n$ is the length of the program.
\end{tsyntax}

\medskip For example, suppose the current goal is
  \ecinput{examps/parts/tactics/swap/1-1.0.ec}
  Then running
  \ecinput{examps/parts/tactics/swap/1-1.ec}
  produces goal
  \ecinput{examps/parts/tactics/swap/1-1.1.ec}
  From which running
  \ecinput{examps/parts/tactics/swap/1-2.ec}
  produces goal
  \ecinput{examps/parts/tactics/swap/1-2.1.ec}
  From which running
  \ecinput{examps/parts/tactics/swap/1-3.ec}
  produces goal
  \ecinput{examps/parts/tactics/swap/1-3.1.ec}
  From which running
  \ecinput{examps/parts/tactics/swap/1-4.ec}
  produces goal
  \ecinput{examps/parts/tactics/swap/1-4.1.ec}
  From which running
  \ecinput{examps/parts/tactics/swap/1-5.ec}
  produces goal
  \ecinput{examps/parts/tactics/swap/1-5.1.ec}
\end{tactic}
